---
title: "HW2 Peer Assessment"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

The Motor Trend Car Road Tests data set was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models). Here, we will perform some analysis on this data set to understand better the effect of hp, wt and disp on mpg.

## Data Description

The data consists of a data frame with 32 observations on the following 11 variables:

1. mpg: Miles/(US) gallon
1. cyl: Number of cylinders
1. disp: Displacement (cu.in.)
1. hp: Gross horsepower
1. drat: Rear axle ratio
1. wt: Weight (1000 lbs)
1. qsec: 1/4 mile time
1. vs: Engine (0 = V-shaped, 1 = straight)
1. am: Transmission (0 = automatic, 1 = manual)
1. gear: Number of forward gears
1. carb: Number of carburetors

We will focus on the the effect of hp, wt and disp on mpg.

## Instructions on reading the data

```{r}
data(mtcars)
```
This loads the data set into your workspace.

# Question 1: Exploratory Data Analysis - 12 points

a. **3 pts** Plot the data (scatterplot) to observe and report the relationship between the response and each of the three predictors hp, wt and disp (there should be 3 plots reported). Comment on the general trend (direction and form).
```{r}
#scatterplot mpg~hp
plot(mpg~hp, data=mtcars) 
abline(lm(mpg~hp, data=mtcars))
```
```{r}
#scatterplot mpg~wt
plot(mpg~wt, data=mtcars) 
abline(lm(mpg~wt, data=mtcars))
```

```{r}
plot(mpg~disp, data=mtcars) 
abline(lm(mpg~disp, data=mtcars))
```

Comment on the general trend: Generally, each plot shows a negative and linear relationship between the response (mpg) and each of the predictor variables (hp, wt and disp for each plot respectively).

b. **3 pts** What is the value of the correlation coefficient for each of the above pair of response and predictor variables? What does it tell you about your comments in part (a).
```{r}
cor(mtcars$mpg, mtcars$hp) 
cor(mtcars$mpg, mtcars$wt) 
cor(mtcars$mpg, mtcars$disp)
```
The correlations between:

mpg and hp is -0.7761684;

mpg and wt is  -0.8676594;

mpg and disp is -0.8475514;

The result shows that there exists a negative correlation between mpg and each of the predictor variables, hp, wt and disp. It shows our hypothesis about the general trend for each plot and the comments that the response is negatively correlated with each of the predictor variables are correct. 

c. **3 pts** Based on this exploratory analysis, is it reasonable to assume a multiple linear regression model for the relationship between mpg and all the predictor variables (hp, wt and disp)? Did you note anything unusual?

Yes, that is a resonable assumption. There is nothing unusual to be seen.

d. **3 pts** Based on the analysis above, would you pursue a transformation of the data?

No, there will not be any transformation because there is a linear relationship between the
response and each of the predictors.


*Please work on non-transformed data for all of the following questions.*

# Question 2: Fitting the Multiple Linear Regression Model - 8 points

Build a multiple linear regression model using the response and all the three predictors and then answer the questions that follow:

a. **4 pts** Report the coefficient of determination for the model and give a single line interpretation of this value.
```{r}
multiLmod <- lm(mpg ~ hp+wt+disp, data=mtcars)
summary(multiLmod)

```

the coefficient of determination for the model is 82.68%.  
Interpretation : 82.68% of the variation in the response is explained by the predictors.

b. **4 pts** Is the model of any use in predicting mpg? Conduct a test of overall adequacy of the model, using $\alpha = 0.05$. Provide the following elements of the test: null hypothesis $H_0$, alternative hypothesis $H_a$, F- statistic or p-value, and conclusion.

$H_0$: $\beta_1$=$\beta_2$=$\beta_3$=0

$H_a$: Not all the $\beta_s$ are equal to 0

The F-test of overall significance of muliple regressing shows the F-statistic is 44.57 on 3 and 28 DF, with a p-value of 8.65e-11. Which reject the null hypothesis.

Conclusion: the result indicates that the muliple linear regression model provides a better fit to the data than a model that contains no independent variables, it will be usefull in predicting mpg.


# Question 3: Checking Assumptions of Model - 15 points

Provide plots to check for Linearity, Constant Variance and Normality assumptions of the model (use your knowledge from Homework 1 Peer Assessment). Provide your interpretations (i.e. whether the assumptions hold) for each plot.

a. **5 pts** Linearity Assumption

**Plot(s):**

For this Linearity assumption, the plots are the same 3 scatterplots as in question1, part (a). The plot are as following:

```{r}
plot(mpg~hp, data=mtcars) 
abline(lm(mpg~hp, data=mtcars))

```

```{r}
plot(mpg~wt, data=mtcars) 
abline(lm(mpg~wt, data=mtcars))

```


```{r}
plot(mpg~disp, data=mtcars) 
abline(lm(mpg~disp, data=mtcars))
```


**Interpretation:**

By the plots, the Linearity assumption holds.

b. **5 pts** Constant Variance Assumption

**Plot(s):**
```{r}
plot(fitted(multiLmod),resid(multiLmod))
abline(h=0)
```


**Interpretation:**

By the plot, the residuals are randomly scattered across zero line, we can assume that the Constant variance assumption holds.

c. **5 pts** Normality Assumption

**Plot(s):**

```{r}
qqnorm(resid(multiLmod))
qqline(resid(multiLmod))
```


**Interpretation:**

By the plot, the Normality assumption holds (except for the rightmost point).

# Question 4: Coefficient Interpretation - 6 points

a. **3 pts** Interpret the coefficient of wt (mention any assumption you make about other predictors clearly when stating the interpretation).

The coefficient of wt is = -3.800891  which means that keeping other predictors constant,  with an increase of 1 unit for wt, there is an decrease in mpg by 3.800891 units .

b. **3 pts** If value of predictor wt in the above model is increased by 0.01 keeping other predictors constant, what change in the response would be expected?

Since the coefficient of wt is -3.800891in the regression model, keeping other predictors constant, with an increase in 0.01, the value of mpg will decrease by = 3.800891  X 0.01 = 0.038 units

 
# Question 4: Confidence Intervals and Interpretation - 9 points

a. **4 pts** Compute 90% and 95% confidence intervals (CIs) for the parameter associated with disp ($\beta_3$) for the model in Question 2.

For the 90% CI:
```{r}
confint(multiLmod, "disp", level = 0.9)
```


For the 95% CI:
```{r}
confint(multiLmod, "disp", level = 0.95)
```

So the 90% CI for disp is [-0.01854328, 0.01666926], the 95% CI for disp is [-0.0221375, 0.02026348].

b. **5 pts** Using just these intervals, what could you deduce about the range (Upper Bound or Lower Bound or both) of the p-value for testing $H_0$: $\beta_3$ = 0 in the model in Question 2?

Since both the CIs include 0, it means that disp is not statistically significant at both levels. Hence, p-value for disp is > 0.05. We can not reject the $H_0$: $\beta_3$ = 0.    Thus, lower bound for p-value is 0.05(exclusive), upper bound is infinity.
